On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-01 tt_theorem.tex
12/1

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be a deterministic unary oritatami system at delay $\delta = 1$, arity $\alpha =2$. 
Assume $\Xi$ stabilizes the transcript until $w[i-1]$. If $S[i] = b$ and $S[i+1] \in \{ t_s, t_o \} $, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

\begin{proof}[lemma \ref{TTT_entrance_Tab}]
Fig.\ref{TTT_tunnel_direction} exhibits all the three kinds of entrance of tunnel A, B.
Let $w[i]$ be stabilized at an entrance point of Tunnel $A$ or $B$.
All cases are $\#bc(C_{i-1}) > \#bc(C_{i})$ as follows.

\begin{itemize}
\item{Case of $t_0$}\\
  Let us consider points $n_3,n_4$. At least one of the points $n_3$ or $n_4$ is free because if both of them are occupied, $p^\prime$ is inside of tunnel. If $n_3$ is free, then $p^\prime$ has to be bound to a bead other than $n_1$ to deterministically stabilize. In this situation, at least three neighbors of $n_1$ are free, that is, $n_1$ has at least one free hand from Lemma~ \ref{TTT_neighbor_lemma}. Hence, $p^\prime$ must be bound to $n_1$. Thus, a case of $t_0$ consumes two hands and it does not supply any binding capabilities.

\item{Case of $t_{\pm 60}$}\\
  In this case, too, $n_4$ or $n_5$ is free. If $n_5$ is free, $p^\prime$ has to be bound to $n_1$ or $n_2$. If $n_5$ is occupied, then $n_4$ is free. This time, by $n_2$ has some free hands so $p^\prime$ has to be bound to $n_2$.

  In this situation, $p^\prime$ is able to supply a binding capabilities which could bind a bead into $n_4$ or $n_5$. However, $n_2$ and $n_3$ are part of a contiguous conformation. According to Jordan curve theorem, any successors of $p^\prime$ cannot reach a point $n_4$ or $n_5$ so this capability is inactive. Thus, in the case of $t_{\pm 60}$ $\#bc(C_{i-1}) > \#bc(C_{i})$.

\item{Case of $t_{\pm 120}$}\\
  Binding capabilities that $p^\prime$ supplies are inactive according to Jordan curve theorem on $n_1$ and $n_2$. Moreover, $p^\prime$ has to be bound to one of $n_3, n_4, n_5$ in order to deterministically stabilize.
Thus, in the case of $t_{\pm 120}$ is $\#bc(C_{i-1}) > \#bc(C_{i})$.

@@

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
	 \draw[->,white] (120:0.8)--(120:0.1);%for adjustment

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

          \node at (0,-2) {$t_{0}$};
        \end{tikzpicture}
      \end{minipage}

            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

          \node at (0,-2) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
            \fill[blue](0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};

@@

            \node[transform canvas={shift=(0:1)},below] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

          \node at (0,-2) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

@@

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$w[i-1]$};

@@

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}
          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (0:1) {$p^\prime$};

            \foreach \theta in {0,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {60,-60,120,-120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel $S$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (-60:1) {$p^\prime$};

            \foreach \theta in {-60,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {0,60,120,-120}{
 	            \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel $O$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
        \begin{tikzpicture}

          \begin{scope}[xshift=2cm, yshift=2cm]
            \draw(0,0) circle [radius=0.05];
            \node[above] at (180:1) {$p^\prime$};
            \node[above] at (-120:1) {$p^\prime$};

            \foreach \theta in {-120,180}{
              \draw[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.05];
            }

            \foreach \theta in {0,60,-60,120}{
              \fill[transform canvas={shift=(\theta:1)}](0,0) circle [radius=0.1];
            }
          \end{scope}

          \node at (2,0) {Tunnel $A$};
        \end{tikzpicture}
      \end{minipage}

    \end{tabular}
    \caption{All possible tunnel sections: straight, obtuse turn, and acute turn}
    \label{TTT_tunnel}
  \end{center}
\end{figure}

@@

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only finite structures whose size is $\mathcal{O}(n)$.
\end{theorem}

@@

\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only finite structures whose size is $\mathcal{O}(n^2)$.
\end{theorem}

@@

\begin{proof}
According to lemma\ref{TTT_a4_2b_lemma}, when $w[i]$ is stabilized, it has to use at least two bonds. Let us consider when a bead $w[i]$ which is the first bead out of $\hexagon_{w[-n+1]}^n$ is stabilized. See Fig.\ref{TTT_a4_first}. any $n_0, n_1, n_3, n_5$ is free because if some of them is occupied, $w[i]$ is not the first bead out of $\hexagon_{w[-n+1]}^n$. At least two neighbors of $w[i]$ except predecessor have to be occupied in order to bind. In this case, a point which is able to put a bead is only $n_2$. Therefore, any transcript cannot be stabilized in out of $\hexagon_{w[-n+1]}^n$. Hence oritatami system can yield only a finite structure whose size is $\mathcal{O}(n^2)$ in $\delta =1, \alpha = 4$.
%lemma???a=4??#bc(C_i) >= #bc(C_{i+1})? ?upper?left
\end{proof}

>>>>

\begin{lemma}
\label{TTT_entrance_Tab}
Let $\Xi$ be a deterministic unary oritatami system at delay $\delta = 1$, arity $\alpha =2$. 
Assume $\Xi$ stabilizes the transcript until $w[i-1]$. If $S[i+1] = b$ and $S[i+2] \in \{ t_s, t_o \} $, then $\#bc(C_{i-1}) > \#bc(C_{i})$.
\end{lemma}

\begin{proof}[lemma \ref{TTT_entrance_Tab}]
See Fig.\ref{TTT_tunnel_direction}.
$S[i+2] \in \{ t_s, t_o \} $ means that $w[i+2]$ is stabilized by a tunnel section type of $S$ or $O$.
Thus,  its predecessor $w[i+1]$ must be inside the tunnel section, that is, $n_1$ and $n_2$ must be occupied.
Free bonds of $w[i]$, if any, cannot be used in future by another bead $w[j]$ because otherwise the part of transcript $w[i..j]$ and the bond between $w[i]$ and $w[j]$ would form a closed curve and the curve would cross the path of $C_{i-1}$ between $n_1$ and $n_2$, contradiction.
%Binding capabilities that $w[i]$ supplies are inactive according to Jordan curve theorem on $n_1$, $n_2$, and $w[i]$.
Therefore, if $w[i]$ forms a bond at its stabilization $\#bc(C_{i-1}) > \#bc(C_{i})$ holds.
Hence, assume $w[i]$ does not form any bond at stabilization.
Then $w[i-1]$ is not on $n_1$ or $n_2$.

On the other hand, let us consider $S[i] = t$.
Beads which are shared by neighbors of $w[i-1]$ and $w[i]$ are occupied so that we consider only Fig.~\ref{TTT_tunnel_direction} (Left) and (Middle) case because (Right) case is not occupied them.
In the (Left) case, if $S[i] = t$, points $n_3$ and $n_4$ are occupied. Then, $S[i+1] = t$, contradiction.
In the (Middle) case, if $S[i] = t$, $n_4$ is free and $n_5$ is occupied because of $S[i+1] = b$.
This time, by $n_4$ has some free hands because at least three neighbors of $n_4$ are free so $w[i]$ has to be bound to $n_2$. Thus, $\#bc(C_{i-1}) > \#bc(C_{i})$.

%In the case of $t_0$, one of $n_3$ and $n_4$ is free because of $S[i+1] = b$. Thus, this case is only $S[i] = b$.
%
%
%In the case of $t_{\pm 60}$, if $S[i] = t$, the point $n_4$ is free and $n_5$ is occupied. This time, by $n_2$ has some free hands because at least three neighbors of $n_2$ are free so $w[i]$ has to be bound to $n_2$. Thus, $\#bc(C_{i-1}) > \#bc(C_{i})$.
%
%In the case of $t_{\pm120}$, the point of $w[i+1]$ is free before $w[i+1]$ is stabilized. Thus, this case is only $S[i] = b$.
%

@@

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
        \draw[->, white] (0:0)--(-120:1);
	 \draw[->,white] (120:0.8)--(120:0.1);%for adjustment

            \fill[blue] (0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

@@

         % \node at (0,-2) {$t_{0}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
	\draw[->, white] (0:0)--(-120:1);
            \fill(0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

@@

          %\node at (0,-2) {$t_{\pm 60}$};
        \end{tikzpicture}
      \end{minipage}

      \begin{minipage}{0.3\hsize}
      \centering
        \begin{tikzpicture}
        \draw[->, white] (0:0)--(-120:1);
            \fill[blue](0,0) circle [radius=0.05];
            \node[below] at (0,0) {$w[i]$};
            \node[below] at (0:1) {$w[i+1]$};

@@

            %\node[transform canvas={shift=(0:1)},below] {$n_0$};
            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},left] {$n_5$};

         % \node at (0,-2) {$t_{\pm 120}$};
        \end{tikzpicture}
      \end{minipage}

@@

            \node[transform canvas={shift=(60:1)},right] {$n_1$};
            \node[transform canvas={shift=(-60:1)},right] {$n_2$};
            \node[transform canvas={shift=(120:1)},left] {$n_3$};
            \node[transform canvas={shift=(-120:1)},left] {$n_4$};
            \node[transform canvas={shift=(180:1)},above] {$w[i-1]$};

@@

@@

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only finite structures whose size is $O(n)$.
\end{theorem}

@@

\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only finite structures whose size is $O(n^2)$.
\end{theorem}

@@

\begin{proof}
According to lemma\ref{TTT_a4_2b_lemma}, when $w[i]$ is stabilized, it has to use at least two bonds. Let us consider when a bead $w[i]$ which is the first bead out of $\hexagon_{w[-n+1]}^n$ is stabilized. See Fig.\ref{TTT_a4_first}. any $n_0, n_1, n_3, n_5$ is free because if some of them is occupied, $w[i]$ is not the first bead out of $\hexagon_{w[-n+1]}^n$. At least two neighbors of $w[i]$ except predecessor have to be occupied in order to bind. In this case, a point which is able to put a bead is only $n_2$. Therefore, any transcript cannot be stabilized in out of $\hexagon_{w[-n+1]}^n$. Hence oritatami system can yield only a finite structure whose size is $O(n^2)$ in $\delta =1, \alpha = 4$.
%lemma???a=4??#bc(C_i) >= #bc(C_{i+1})? ?upper?left
\end{proof}
